---
title: "Project 2: Data Tidying"
author: "Lucy Watts, Natalie Lau, Catherine Weeks, and Nandir Altansukh"
format: html
---

## Data Tidying

```{r}
#| warning: false
#| message: false
#| echo: false

library(tidyverse)
library(lubridate)

spotify_songs_raw <- read.csv("spotify_songs.csv")
```


We can see that there are some repeat tracks in this data set, since the number of observations is greater than the number of unique track ids.

```{r}
paste("Number of Tracks: ", nrow(spotify_songs_raw))
paste("Number of Unique Tracks: ", length(unique(spotify_songs_raw$track_id)))
```
 
 
Exploratory analysis part! To my understanding we’re trying to discover possible relationships and trends
 
 
 

### Data Wrangling

```{r}
# Data wrangling

spotify_songs_cleaned <- spotify_songs_raw |>
  mutate(track_album_release_date = ymd(track_album_release_date)) |>
  select(track_id, track_name, track_artist, track_popularity, track_album_release_date, danceability, energy, key,
         loudness, mode, valence, tempo, duration_ms) |>
  filter(year(track_album_release_date) == 2019) |>
  filter(str_detect(track_name,
                    "Remix|Remaster|Remastered|REMIX|remix") == FALSE) |>
  unique() |>
  mutate(song_percentile = percent_rank(track_popularity)*100) |>
  filter(song_percentile >= 90) |>
  arrange(desc(song_percentile))

```

Steps taken in data wrangling:

- changed the variable `track_album_release_date` to a `Date` object
- removed the variables that weren't of interest to our project
- filtered the data to only tracks from 2019
- removed all remixes and remastered versions of tracks
- removed all duplicate tracks
- calculated the popularity of each track
- filtered the data to only the top 10% of tracks

NOTE: There are still some duplicate songs, songs that were released first as a single and then as part of an album, and therefore have different track ids.

```{r}
write_csv(spotify_songs_cleaned, "spotify_cleaned.csv")
```

