---
title: "Project 2: Overall Analysis"
author: "Lucy Watts, Natalie Lau, Catherine Weeks, and Nandir Altansukh"
format: html
---

## Overall Data Analysis

```{r}
#| warning: false
#| message: false
#| echo: false

# Set Up
library(tidyverse)
library(lubridate)

spotify_cleaned <- read_csv("spotify_cleaned.csv")
```

Variables for Analysis:

- Energy
- Loudness
- Tempo
- Danceability


### Energy

```{r}
energy_histogram <- spotify_cleaned |>
  ggplot(aes(x = energy)) +
  geom_histogram(binwidth = .05, color = "white") +
  theme_minimal() +
  labs(title = "Energy vs Track Popularity (Top 10% of Most Popular 2019 Spotify Songs)",
       x = "Energy",
       y = "Track Popularity")
energy_histogram
```


### Loudness

```{r}
loudness_plot <- ggplot(spotify_cleaned,
       aes(x = loudness, y = track_popularity)) +
  geom_point(alpha = 0.6) +
  theme_minimal() +
  labs(title = "Loudness and Track Popularity",
       x = "Loudness",
       y = "Track Popularity")

loudness_plot
```
Track popularity seems to rise for loudness of song ranges between -13 to -4, as track popularity rises above 90. Songs with loudness rating outside of that range seems to have track popularities below 90, indicating that there is a correlation between loudness of a song and how many streams it receives.


### Tempo

``` {r}
tempo_plot <- spotify_cleaned |>
  ggplot(aes(x = tempo, y = track_popularity)) +
  geom_point(position = "jitter", alpha = 0.3) +
  geom_smooth(method = "loess", se = FALSE, color = "red") +
  labs(title = "Tempo vs Popularity in the Top 10% Most Popular Spotify Songs, 2019",
       x = "Tempo (bpm)",
       y = "Popularity Rating (0-100)") +
  theme_minimal()

tempo_plot
```


### Danceability

```{r}
mean_dance = spotify_cleaned |>
  summarize(mean_dance = mean(danceability)) |>
  pull()

dance_plot <- spotify_cleaned |>
  ggplot(aes(x = danceability, y = track_popularity)) +
  geom_point(position = "jitter") +
  geom_vline(xintercept = mean_dance, color = "red") +
  labs(
  title = "Danceability vs Track Popularity in the Top 10% Most Popular Spotify Songs, 2019",
  x = "Danceability",
  y = "Track Popularity"
  ) + 
  theme_minimal()

dance_plot
```

