---
title: "Project 2: Overall Analysis"
author: "Lucy Watts, Natalie Lau, Catherine Weeks, and Nandir Altansukh"
format: html
---

## Overall Data Analysis

```{r}
#| warning: false
#| message: false
#| echo: false

# Set Up
library(tidyverse)
library(lubridate)

spotify_songs_cleaned <- read_csv("spotify_cleaned.csv")
```

Variables for Analysis:

- Energy
- Loudness
- Tempo
- Danceability


### Energy

```{r}

```


### Loudness


```{r}
ggplot(spotify_songs_cleaned,
       aes(x = loudness, y = track_popularity)) +
  geom_point(alpha = 0.6) +
  theme_minimal() +
  labs(title = "Loudness and Track Popularity",
       x = "Loudness",
       y = "Track Popularity")

```
Track popularity seems to rise for loudness of song ranges between -13 to -4, as track popularity rises above 90. Songs with loudness rating outside of that range seems to have track popularities below 90, indicating that there is a correlation between loudness of a song and how many streams it receives.


### Tempo

```{r}

```


### Danceability

```{r}

```

=======
